State of the Practice for Lattice Boltzmann Method Software

### 2022-04-25 StateOfPractice/Papers/LatticeBoltz_SOP-arXiv/LatticeBoltz_SOP.tex
Revisions to Section 7

Twenty-three packages use issue tracking, 15 of which use GitHub or GitLab to
host their git repository, seven use email, and one (SunlightLB) uses
SourceForge. Most software packages that use GitHub or GitLab have the majority
of their issues closed, and only three (laboetie, lettuce, Sailfish) have less
than 50\% of their issues closed. All the top five ranked packages have most of
their issues closed. Alive packages (11 use issue tracking) have 64\% of their
issues closed, while dead packages (3 use issue tracking) have 71\% of their
issues closed. Table~\ref{gitrepodata} presents this information. With respect
to the host for the git repository, 13 packages use GitHub and two (Palabos,
waLBerla) use GitLab. Of the other packages, one package (SunlightLB) uses CVS
for issue tracking and version control, and seven do not appear to use any issue
tracking system.

@@

Developers use software tools to support the development, verification,
maintenance, and evolution of software, software processes, and artifacts
\citep[p.\ 501]{GhezziEtAl2003}. Table~\ref{tbl_tools} summarizes the tools,
subdivided into development tools, dependencies, and project management tools.
In this section we answer \rqref{RQ_CompareToolsProjMngmnt} by comparing aspects
of tool usage in LBM software packages to their use in the research software
community in general.

@@

Development tools support the development of end products, but do not become
part of them, unlike dependencies that remain in the application once it is
released \citep[p.\ 506]{GhezziEtAl2003}. Although not shown in
Table~\ref{tbl_tools}, developers also likely use debuggers.  Only three
(ESPResSo, Ludwig and Musubi) packages mentioned CI tools, like Travis CI.
Several packages explicitly noted code editors and compilers, but likely all
packages use them. One of the packages (Ludwig) explicitly noted the use of
proprietary unit testing code written in C. Likewise, one of the developers
(pyLBM) noted the use of proprietary code for verifying the correctness of
output. Developers of other packages likely used similar tools.

@@

Many of the software packages that were assessed were developed by teams of two
or more people. Their work needed to be coordinated and managed.
Table~\ref{tbl_tools} shows the types of project management tools that were
explicitly noted in the artifacts, webpages, or interviews with the developers.
As with development tools and dependencies, developers may use other types of
project management tools, but they were not visible in the artifacts to which we
had access.  Collaboration tools, most often email and video, are noted as being
used when developing software projects. Our interviewees did not mention project
management software, but the sample is not large enough to rule out other
projects using such software. GitHub is the home for many of the projects.
Developers use the platform to help manage their projects, especially bug
related issues. Most of the projects appear to use change tracking and version
control tools. Half of the packages mention document generation tools, like
Sphinx and Doxygen.

Based on information provided by \citet{JungEtAl2022}, tool usage for
LBM software has much in common with tool usage for ocean modelling
software.  Both use tools for editing, compiling, code management, testing,
building, and project management.  From the data available, ocean modelling
differs from LBM software in its use of Kanban boards for project management.
LBM software difference from the ocean domain through the use of continuous
integration, and document generation tools.

@@

The poor adoption of version control tools that Wilson lamented in 2006
\citep{Wilson2006} has greatly improved in the intervening years.  From
Section~\ref{Sec_Maintainability}, 67\% of LBM packages use version control
(GitHub, GitLab, or CVS).  (From Table~\ref{gitrepodata}, 11/15 or 73\% of alive
packages use version control.)  The proliferation of version control tools for
LBM matches the increase in the broader research software community.
\citet{Nguyen-HoanEtAl2010} estimated a little over 10 years ago that only 50\%
of research software projects use version control, but even at that time
\citet{Nguyen-HoanEtAl2010} noted an increase from previous usage levels. A
survey in 2018 shows 81\% of developers use a version control system
\citep{AlNoamanyAndBorghi2018}.  \citet{Smith2018} has similar results, showing
version control usage for alive projects in mesh generation, geographic
information systems and statistical software for psychiatry increasing from
75\%, 89\% and 17\% (respectively) to 100\%, 95\% and 100\% (respectively) over
a 4-year period ending in 2018.  (For completeness the same study showed a
decrease in version control usage for seismology software over the same time
period, from 41\% down to 36\%).  Almost every software guide cited in
Section~\ref{Sec_CompareArtifacts} includes the advice to use version control.
The high usage of version control tools in LBM software matches the trend in
research software in general.

@@

As mentioned in Section~\ref{SecSurfCorrectAndVerifiab}, LBM developers rarely
uses CI (3 of 24 packages or 12.5\%), although interviews with developers
suggest that the actual rate might be higher. Low use of CI for LBM contrasts
with how frequently it is recommended in research software development
guidelines
\citep{BrettEtAl2021, Brown2015, ThielEtAl2020, Zadka2018, vanGompelEtAl2016}
and its popularity with non-research software. \citet{HiltonEtAl2016} show CI
usage across a large sample of GitHub projects (34,544 projects) is at least
40\%, with more popular projects (as measured by stars), having even higher
usage rates. Moreover, \citet{HiltonEtAl2016} predict that CI adoption rates for
open-source projects will increase even further in the future.  We could not
find published data on the CI usage rates for research software in general.
However, we saw 17\% utilization in the specific domain of imaging software
\citep{Dong2021}, which is comparable to the LBM usage rate. Our impression is
that medical imaging software and LBM software are not unique --- most research
software domains likely lag behind the best practice recommendations, and
non-research software usage rates, for continuous integration.

CI is the process of building and testing software on every push to the code
repository, with pushes done very frequently \citep[p.\ 13]
{HumbleAndFarley2010}, \citep{ShahinEtAl2017, Fowler2006}.  CI consists of the
following elements:
\begin{itemize}
	\item A version control system \citep{Fowler2006}. To be effective, all
	files should be under version control, not just code files.  Anything that
	is needed to build, install and run the software should be under version
	control, including configuration files, build scripts, test harnesses, and
	operating system configurations \citep[p. 19]{HumbleAndFarley2010}.
	Fortunately for the LBM domain, as discussed above, most alive projects
	(73\% of those assessed) use version control.
	\item A fully automated build system \citep{Fowler2006}.  As \citet[p.\
	5]{HumbleAndFarley2010} point out, deploying software manually is an
	anti-pattern.  Fortunately, as discussed above, most LBM projects use build
	automation.
	\item An automated test system \citep{Fowler2006}. Building quality software
	involves creating automated tests at the unit, component, and acceptance
	test level, and executing these tests whenever someone makes a change to the
	code, its configuration, the environment, or the software stack that it runs
	on \citep[p.\ 83]{HumbleAndFarley2010}. As Table~\ref{artifactspresent}
	shows, test cases are in the uncommon category for LBM software artifacts,
	which means that some LBM projects will need to increase their testing
	automation if they wish to pursue CI.  The usual advice for CI is to keep
	the build and test process short \citep[p.\ 60]{HumbleAndFarley2010}. Given
	that LBM can be computationally expensive, the tests run with every check-in
	may need to focus on simple code interface tests, saving larger more complex
	tests for less frequent runs.  A more sophisticated option for when tests
	create a bottleneck for merges is CIVET (Continuous Integration,
	Verification, Enhancement, and Testing), which solves this problem by
	intelligently pinning, cancelling, and if necessary, restarting jobs as
	merges occur, so that two merges are being tested at all times
	\citep{SlaughterEtAl2021}.
	\item An automated system for other tasks, such as code checking,
	documentation building and web-site updating.  These other tasks are not
	essential parts of CI, but they can be incorporated to improve the quality
	of the code and the communication between developers and users. For
	instance, a static analysis (possibly via linters) of the code may find poor
	programming practice or lack of adherence to adopted coding standards.
	\citet{SlaughterEtAl2021} provides another example of an automated task ---
	checking that a test specification includes the test's motivation, a test
	description, and a design description for all changes. 
	\item An integration build system to pull everything together.  Every time
	there is a check-in (for instance a pull request), the integration server
	automatically checks out the sources onto the integration machine, starts a
	build, runs tests, and informs the committer of the results. 
\end{itemize}

@@

Although there are still challenges, setting up a CI system has never been
easier than it is today.  The option of installing a dedicated CI server (either
physically or virtually) exists with tools such as
\href{https://www.jenkins.io/} {Jenkins}, \href{http://buildbot.net/}
{Buildbot}, \href{https://www.gocd.org/} {Go}, and
\href{http://integrity.github.io/} {Integrity}. However, installation on your
own server is not necessary since there are many hosted CI tool solutions
available, such as: \href{https://travis-ci.org/} {Travis CI},
\href{https://github.com/features/actions} {GitHub Actions} and
\href{https://circleci.com/} {CircleCI}.  Getting started with a hosted CI is
straightforward.  All that is required to begin is editing a few lines of a
simple YAML configuration file in the project's root directory.

@@

We answer research question \rqref{RQ_CompareMethodologies} by comparing the
principles, processes, and methodologies used for LBM software to what can be
gleaned from the literature on research software in general. Our measured data
on LBM software artifacts does not explicitly indicate the development process.
However, during our interviews one developer (ESPResSo) told us their
non-rigorous development model is like a combination of agile and waterfall.
Employing a loosely defined process makes sense for LBM software, given that the
teams are generally small and self-contained.  Although eleven of the packages
explicitly convey that they would accept outside contributors, teams are
generally centralized, often working at the same institution. Working at the
same institution means that an informal process can show success, since informal
conversations are relatively easy to have.

Interviews with developers confirmed a similar project management process. In
teams of only a couple of developers, the entire team discusses additions of new
features or major changes. Projects with more than a couple developers have lead
developer roles. These lead developers review potential additions to the
software. One of the developers (ESPResSo) noted that they use an ad hoc peer
review process to assess major changes and additions. Using peer review (also
called technical review) matches with recommended practice for research software
\citep{HerouxEtAl2008, Givler2020, OrvizEtAl2017, USGS2019}.

@@

Most of the software packages do not explicitly state the motivations or design
principles that were considered when developing the software. One package,
Sailfish, indicates in its artifacts the explicit goal of shortening development
time, with the developers using Python and CUDA/OpenCL to achieve this without
sacrificing performance. The Sailfish documentation explicitly lists the goals
of performance, scalability, agility and extendability, maintenance, and ease of
use. The project scored well in these categories during our assessment.  The
quality priorities for Sailfish roughly match the priorities observed for
research software in general. \citet{Nguyen-HoanEtAl2010} surveyed developers to
find the following list of qualities, in decreasing order of importance:
reliability, functionality, maintainability, availability, performance,
flexibility, testability, usability, reusability, traceability, and portability.
The Sailfish list does not list reliability or functionality, but we can safely
assume those are implicitly high priorities for any scientific project.  In
earlier studies \citet{KellyAndSanders2008} and \citet{CarverEtAl2007} highlight
how important correctness is for research software.

@@

During our interviews, developers noted that documentation plays a significant
role in the development process, specifically with on-boarding new developers. A
goal of documentation is to lower the entry barrier for these new contributors.
The documentation provides information on how to get started, orients the user
to artifacts and the source code, and explains how the system works, including
the so-called simulation engine and interface.  This emphasis on documentation,
especially for new developers, is echoed in research software guidelines.
Multiple guidelines recommend a document explaining how to contribute to a
project, often named CONTRIBUTING \citep{Yo2021, BrettEtAl2021, WilsonEtAl2016,
ThielEtAl2020, vanGompelEtAl2016, OrvizEtAl2017, FLOSS2022, JimenezEtAl2017}.
Tutorials \citep{ThielEtAl2020}, troubleshooting guides \citep{OrvizEtAl2017,
SSI2022} and quick start examples \citep{ThielEtAl2020, vanGompelEtAl2016} are
also recommended.  \citet{SmithAndRoscoe2018} suggests including instructions
specifically for on-boarding new developers. For open source software in general
(not just research software), \citet{Fogel2005} recommends providing tutorial
style examples, developer guidelines, demos, and screenshots.

@@

\begin{itemize}
	\item Explicitly state the design goals and priorities, so that those using
	and developing your software know part of the rationale for the design
	decisions. Sailfish is an example where this has been done, as discussed in
	Section~\ref{Sec_DesPrincipAndPriorities}.
	\item Write down the likely changes that the design is intended to support.
	Since most designs are implicitly (or explicitly) motivated to facilitate
	future change, recording the likely changes can help readers understand the
	design.
	\item Explicitly record the philosophy or principle behind the
	decomposition. One potential principle, related to the previously mentioned
	list of likely changes, is the principle of information hiding
	\citep{Parnas1972a}.  This principle supports design for change through the
	``secrets'' of each module. A text blurb can be given for each module or
	class to describe its secrets and the services it provides.
	\item Rigorously document the design following the template for a Module
	Guide (MG) \citep{ParnasEtAl1984}.  An MG organizes the modules in a
	hierarchy by their secrets. The application of the Parnas approach to
	scientific software has been illustrated by applying it to the example of a
	mesh generator \citep{SmithAndYu2009}.
	\item Graphically represent the design using Unified Modelling Language
	class diagrams.  This approach is suited to object-oriented design and
	designs that use patterns \citep{Gamma1995}.
	\item Explain the design using data flow diagrams to show typical use cases
	for input transformation.
	\item Using a textual description, list for each module (or class), the
	state variables (if any), exported constants and all exported access
	programs.  This shows the interface that can used to access each module's
	services.
	\item Document the Module Interface Specification (MIS)
	\citep{HoffmanAndStrooper1995}. An MIS is an abstract model that 
	formally shows the state variables, inputs and outputs
	\citep{ElSheikhEtAl2004, SmithAndYu2009}.
\end{itemize}

@@

This section summarizes the best practices from LBM developers that are taken to
address the pain points mentioned in Section~\ref{painpoints}.  The main source
of information is the qualitative data from developer interviews
(Section~\ref{SecSurvey}).  The practices summarized in this section can
potentially be emulated by the LBM software packages that do not currently
follow them.  Moreover, these practices may also provide examples that can be
followed by other research software domains.

>>>>

Twenty-three packages use issue tracking, 15 of which use GitHub or GitLab to
host their git repository, seven use email, and one (SunlightLB) uses
SourceForge. Most software packages that use GitHub or GitLab have the majority
of their issues closed, and only three (laboetie, lettuce, Sailfish) have less
than 50\% of their issues closed. All the top five ranked packages have most of
their issues closed. Alive packages (11 use issue tracking) have 64\% of their
issues closed, while dead packages (3 use issue tracking) have 71\% of their
issues closed. Table~\ref{gitrepodata} presents this information. With respect
to the host for the git repository, 13 packages use GitHub and two (Palabos,
waLBerla) use GitLab. Of the other packages, one package (SunlightLB) uses CVS
for issue tracking and version control, and seven do not appear to use any issue
tracking system.  Assuming the use of GitHub, GitLab or CVS implies the use of
version control, 16 of 24 (67\%) of projects use version control.

@@

Developers use software tools to support the development, verification,
maintenance, and evolution of software, software processes, and artifacts
\citep[p.\ 501]{GhezziEtAl2003}. Table~\ref{tbl_tools} summarizes the observed
tools, subdivided into development tools, dependencies, and project management
tools. In this section we answer \rqref{RQ_CompareToolsProjMngmnt} by comparing
aspects of tool usage in LBM software packages to their use in the research
software community in general.

@@

Development tools support the development of end products, but do not become
part of them, unlike dependencies that remain in the application once it is
released \citep[p.\ 506]{GhezziEtAl2003}. Although not shown in
Table~\ref{tbl_tools}, developers also likely use debuggers.  Only three
(ESPResSo, Ludwig and Musubi) packages mention CI tools, like Travis CI. Several
packages explicitly note code editors and compilers, but likely all packages use
them. One of the packages (Ludwig) explicitly notes the use of proprietary unit
testing code written in C. Likewise, one of the developers of pyLBM mentions the
use of proprietary code for verifying the correctness of output. Developers of
other packages likely use similar tools.

@@

Many of the software packages have teams of two or more. Their work needs to be
coordinated and managed. Table~\ref{tbl_tools} shows the types of project
management tools that were explicitly noted in the artifacts, webpages, or
interviews with the developers. As with development tools and dependencies,
developers may use other types of project management tools, but they were not
visible in the artifacts to which we had access.  For collaboration tools the
most often observed are email and video. Our interviewees did not mention
project management software, but our sample is not large enough to rule out
other projects using such software. GitHub is the home for many of the projects.
Developers use the platform to help manage their projects, especially bug
related issues. Most of the projects appear to use change tracking and version
control tools. Half of the packages mention document generation tools, like
Sphinx and Doxygen.

Based on information provided by \citet{JungEtAl2022}, tool usage for LBM
software has much in common with tool usage for ocean modelling software.  Both
use tools for editing, compiling, code management, testing, building, and
project management.  From the data available, ocean modelling differs from LBM
in its use of Kanban boards for project management. LBM software differs from
the ocean modelling through the use of continuous integration, and document
generation.

@@

The poor adoption of version control tools that Wilson lamented in 2006
\citep{Wilson2006} has greatly improved in the intervening years.  From
Section~\ref{Sec_Maintainability}, 67\% of LBM packages use version control
(GitHub, GitLab, or CVS). If we restrict our attention to alive packages,
Table~\ref{gitrepodata}, shows 11/15 or 73\% of alive packages use version
control. The proliferation of version control tools for LBM matches the increase
in the broader research software community. \citet{Nguyen-HoanEtAl2010}
estimated a little over 10 years ago that only 50\% of research software
projects use version control, but even at that time they noted an increase from
previous usage levels. A survey in 2018 shows 81\% of developers use a version
control system \citep{AlNoamanyAndBorghi2018}. \citet{Smith2018} has similar
results, showing version control usage for alive projects in mesh generation,
geographic information systems and statistical software for psychiatry
increasing from 75\%, 89\% and 17\% (respectively) to 100\%, 95\% and 100\%
(respectively) over a 4-year period ending in 2018.  (For completeness the same
study showed a decrease in version control usage for seismology software over
the same time period, from 41\% down to 36\%).  Almost every software guide
cited in Section~\ref{Sec_CompareArtifacts} includes the advice to use version
control. The high usage of version control tools in LBM software matches the
trend in research software in general.

@@

As mentioned in Section~\ref{SecSurfCorrectAndVerifiab}, LBM developers rarely
uses CI (3 of 24 packages or 12.5\%), although interviews with developers
suggest that the actual rate might be higher. Low use of CI for LBM contrasts
with how frequently it is recommended in research software development
guidelines
\citep{BrettEtAl2021, Brown2015, ThielEtAl2020, Zadka2018, vanGompelEtAl2016}
and its popularity with non-research software. \citet{HiltonEtAl2016} show CI
usage across a large sample of GitHub projects (34,544 projects) is at least
40\%, with more popular projects (as measured by stars), having rates up to
70\%. Moreover, \citet{HiltonEtAl2016} predict that CI adoption rates for
open-source projects will increase even further in the future.  We could not
find published data on the CI usage rates for research software in general.
However, we saw 17\% utilization in the specific domain of imaging software
\citep{Dong2021}, which is comparable to the LBM usage rate. Our impression is
that medical imaging software and LBM software are not unique --- most research
software domains likely lag behind the best practice recommendations, and
non-research software usage rates, for CI.

Continuous integration is the process of building and testing software on every
push to the code repository, with pushes done very frequently \citep[p.\ 13]
{HumbleAndFarley2010}, \citep{ShahinEtAl2017, Fowler2006}.  CI consists of the
following elements:
\begin{itemize}
	\item A version control system \citep{Fowler2006}. To be effective, all
	files should be under version control, not just code files.  Anything that
	is needed to build, install and run the software should be under version
	control, including configuration files, build scripts, test harnesses, and
	operating system configurations \citep[p. 19]{HumbleAndFarley2010}.
	Fortunately for the LBM domain, as discussed above, most alive projects
	(73\% of those assessed) use version control.
	\item A fully automated build system \citep{Fowler2006}.  As \citet[p.\
	5]{HumbleAndFarley2010} point out, deploying software manually is an
	anti-pattern.  Fortunately, as discussed above, most LBM projects use build
	automation.
	\item An automated test system \citep{Fowler2006}. Building quality software
	involves creating automated tests at the unit, component, and acceptance
	test level, and executing these tests whenever someone makes a change to the
	code, its configuration, the environment, or the software stack that it runs
	on \citep[p.\ 83]{HumbleAndFarley2010}. As Table~\ref{artifactspresent}
	shows, test cases are in the uncommon category for LBM software artifacts,
	which means that some LBM projects will need to increase their testing
	automation if they wish to pursue CI.  The usual advice for CI is to keep
	the build and test process short \citep[p.\ 60]{HumbleAndFarley2010}. Given
	that LBM can be computationally expensive, the tests run with every check-in
	may need to focus on simple code interface tests, saving large tests for
	less frequent runs.  A more sophisticated option to address the bottleneck
	for merges is CIVET (Continuous Integration, Verification, Enhancement, and
	Testing), which solves this problem by intelligently pinning, cancelling,
	and if necessary, restarting jobs as merges occur \citep{SlaughterEtAl2021}.
	\item An automated system for other tasks, such as code checking,
	documentation building and web-site updating.  These other tasks are not
	essential to CI, but they can be incorporated to improve the quality
	of the code and the communication between developers and users. For
	instance, a static analysis (possibly via linters) of the code may find poor
	programming practice or lack of adherence to adopted coding standards.
	\citet{SlaughterEtAl2021} provides another example of an automated task ---
	checking that a test specification includes the test's motivation, a test
	description, and a design description for all changes. 
	\item An integration build system to pull everything together.  Every time
	there is a check-in (for instance a pull request), the integration server
	automatically checks out the sources onto the integration machine, starts a
	build, runs tests, and informs the committer of the results. 
\end{itemize}

@@

Although there are still challenges, setting up a CI system has never been
easier than it is today.  The option of installing a dedicated CI server (either
physically or virtually) exists with tools such as
\href{https://www.jenkins.io/} {Jenkins}, \href{http://buildbot.net/}
{Buildbot}, \href{https://www.gocd.org/} {Go}, and
\href{http://integrity.github.io/} {Integrity}. However, installation on your
own server is unnecessary since there are many hosted CI tool solutions
available, such as: \href{https://travis-ci.org/} {Travis CI},
\href{https://github.com/features/actions} {GitHub Actions} and
\href{https://circleci.com/} {CircleCI}.  Getting started with a hosted CI is
straightforward.  All that is required to begin is editing a few lines of a YAML
configuration file in the project's root directory.

@@

We answer research question \rqref{RQ_CompareMethodologies} by comparing the
principles, processes, and methodologies used for developing LBM software to
those used for research software in general. Our measured data on LBM software
artifacts does not explicitly indicate the development process. However, during
our interviews one developer (ESPResSo) told us their non-rigorous development
model is like a combination of agile and waterfall. Employing a loosely defined
process makes sense for LBM software, given that the teams are generally small
and self-contained.  Although eleven of the packages explicitly convey that they
would accept outside contributors, teams are generally centralized, often
working at the same institution. Working at the same institution means that an
informal process can show success, since informal conversations are relatively
easy to have.

Our interviews with LBM developers confirmed a similar project management
process. In teams of only a couple of developers, the entire team discusses
additions of new features or major changes. Projects with more than a couple
developers have lead developer roles. These lead developers review potential
additions to the software. One of the developers (ESPResSo) noted that they use
an ad hoc peer review process to assess major changes and additions. Using peer
review (also called technical review) matches with recommended practice for
research software \citep{HerouxEtAl2008, Givler2020, OrvizEtAl2017, USGS2019}.

@@

Most of the software packages do not explicitly state the motivations or design
principles that were considered when developing the software. One package,
Sailfish, indicates in its artifacts the explicit goal of shortening development
time, with the developers using Python and CUDA/OpenCL to achieve this without
sacrificing performance. The Sailfish documentation explicitly lists the goals
of performance, scalability, agility and extendability, maintenance, and ease of
use. The project scored well in these categories during our assessment.  The
quality priorities for Sailfish roughly match the priorities observed for
research software in general. \citet{Nguyen-HoanEtAl2010} surveys developers to
find the following list of qualities, in decreasing order of importance:
reliability, functionality, maintainability, availability, performance,
flexibility, testability, usability, reusability, traceability, and portability.
The Sailfish list does not list reliability or functionality, but we can safely
assume those are implicitly high priorities for any scientific project.  In
earlier studies \citet{KellyAndSanders2008} and \citet{CarverEtAl2007} highlight
how important correctness is for research software.

@@

During our interviews, developers noted that documentation plays a significant
role in the development process, specifically with on-boarding new developers. A
goal of documentation is to lower the entry barrier for these new contributors.
The documentation provides information on how to get started, orients the user
to artifacts and the source code, and explains how the system works, including
the so-called simulation engine and interface.  This emphasis on documentation,
especially for new developers, is echoed in research software guidelines
summarized in Table~\ref{Tbl_Guidelines}. For open source software in general
(not just research software), \citet{Fogel2005} recommends providing tutorial
style examples, developer guidelines, demos, and screenshots.

@@

\begin{itemize}
	\item Explicitly state the design goals and priorities, so that those using
	and developing your software know part of the rationale for the design
	decisions. Sailfish is an example where this has been done, as discussed in
	Section~\ref{Sec_DesPrincipAndPriorities}.
	\item Write down the likely changes that the design is intended to support.
	Since most designs are implicitly (or explicitly) motivated to facilitate
	future change, recording the likely changes can help readers understand the
	design.
	\item Explicitly record the philosophy or principle behind the
	decomposition. One potential principle, related to the previously mentioned
	list of likely changes, is the principle of information hiding
	\citep{Parnas1972a}.  This principle supports design for change through the
	``secrets'' of each module. A text blurb can be given for each module or
	class to describe its secrets and the services it provides.
	\item Rigorously document the design following the template for a Module
	Guide (MG) \citep{ParnasEtAl1984}.  An MG organizes the modules in a
	hierarchy by their secrets. The application of the Parnas approach to
	research software has been illustrated by applying it to the example of a
	mesh generator \citep{SmithAndYu2009}.
	\item Graphically represent the design using Unified Modelling Language
	class diagrams.  This approach is suited to object-oriented design and
	designs that use patterns \citep{Gamma1995}.
	\item Explain the design using data flow diagrams to show typical use cases
	for input transformation.
	\item Using a textual description, list for each module (or class), the
	state variables (if any), exported constants and all exported access
	programs.  This shows the interface that can used to access each module's
	services.
	\item Document the Module Interface Specification (MIS)
	\citep{HoffmanAndStrooper1995}. An MIS is an abstract model that formally
	shows each module's access programs and the associated transitions and
	outputs based on their state, environment, and input variables
	\citep{ElSheikhEtAl2004, SmithAndYu2009}.
\end{itemize}

@@

This section answers \rqref{RQ_Concerns} by summarizing the best practices LBM
developers take to address the pain points mentioned in
Section~\ref{painpoints}.  The main source of information is the qualitative
data from developer interviews (Section~\ref{SecSurvey}).  The practices
summarized in this section can potentially be emulated by the LBM software
packages that do not currently follow them.  Moreover, these practices may also
provide examples that can be followed by other research software domains.
